Improving \textsl{Gaia} parallaxes with a data-driven model \\ of the color--magnitude diagram

### 2017-03-16 paperDistance/paper.tex
m67 section

We use stars crossmatched in TGAS, 2MASS and within the observing footprint of \panstarrs to access the Schlafly 3D dust model. We require the signal to noise of both the photometry and parallax be $> 0.001$, that the photometry have nonzero errors, and remove a small selection of 2MASS stars that have zero color and zero J band magnitude. We correct for dust using 3D Bayestar (Schlafly+Finkbeiner) with an estimate of the distance from iteratively sampling our posterior distances, more below.

@@

Dust corrections are non trivial for stars with poor signal to noise. Stars with poor signal to noise have a likelihood that is consistent with infinite distance and can therefore have severe dust corrections. This is most obvious in the giant stars, which tend to have the lowest signal to noise, and can therefore have dust corrections $> 2$ magnitudes. To apply a dust correction to the \tmass photometry, we first generate our prior using \xd\ on the observed, attenuated photometry. Using this raw prior we infer the distances to all the stars in our prior more precisely, increasing our signal to noise for each star and therefore minimize severe distances and their associated severe dust corrections. With more precise distance posteriors, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the Green et al 3D dust map at that distance. For each distance we take the median of the samples of the dust model. We apply each dust correction to our 2MASS photometry given by Equation \ref{eq:dust} where $K_{\lambda}$ = [3.626, 2.742, 3.303, 2.285, 1.698, 0.709, 0.449, 0.302] for bands [B, V, g, r, i, J, H, K] respectively, taken from table (REF TABLE SFD), and $E(B-V)$ is the output of the Green et al dust models. We then regenerate our prior using the dust corrected photometry. We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads again to a lighter deconvolution (Do we want to change this and add in the covariance for the dust?). Should I quote some statistics about the dust corrections? mostly small? maybe difference before after correcting for distance?

\begin{equation}
m_{\mathrm{band, corrected}} = m_{\mathrm{band}} - Q_{\lambda} \times E(B-V)
\label{eq:dust}
\end{equation}

>>>>

We use stars crossmatched in TGAS, 2MASS and within the observing footprint of \panstarrs\ to access the Greene 3D dust model. We require the photometry have real values and nonzero, real errors, and remove a small selection of \tmass\ stars that have zero color and zero J band magnitude.

@@

We correct for dust using 3D Bayestar (Schlafly+Finkbeiner) with an estimate of the distance from iteratively sampling our posterior distances. Dust corrections are non trivial for stars with poor signal to noise. Stars with poor signal to noise have a likelihood that is consistent with infinite distance and can therefore have severe dust corrections. This is most obvious in the giant stars, which tend to have the lowest signal to noise, and can therefore have dust corrections $> 2$ magnitudes. To apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry. Using this raw prior we infer more precise distances to all the stars which we can use to get a measurement of dust. This approach minimizes severe distances from the likelihoods alone and their associated severe dust corrections. With more these more precise distance posteriors, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the Green et al 3D dust map at that distance. For each distance we take the median of the samples of the dust model. We apply each dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (REF TABLE SFD), and $E(B-V)$ is the output of the Green et al dust models. We then regenerate our prior using the dust corrected photometry. We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads again to a lighter deconvolution (Do we want to change this and add in the covariance for the dust?). We iterate this process 10 times, however the dust values seem to converge after about 2-3 iterations. Should I quote some statistics about the dust corrections? mostly small? maybe difference before after correcting for distance?

LMA: what are our results?
\subsection{Distances to M67}
As proof of concept we inferred the distances to stars in the open cluster M67. It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away. Previous work has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5. With \gaia\ parallaxes having a minimum uncertainty of about 0.3 mas, for stars with the most certain parallaxes, the SN reaches 5 at about 700 pc, right where this cluster lies. So it is a great test case for comparing parallax inferences with or without a prior. We made a window of 1 degree squared on the sky with the cluster M67 lying in the middle. The cluster is visibly obvious as a clustering of stars in a visualization of the \gaia\ data in 2D on the sky. With our converged dust values, and it's associated prior, we calculate the posterior of the true parallaxes to the stars within this window on the sky around M67. Figure XXX shows the comparison of the likelihoods and posteriors. The vertical lines bracket the parallaxes of, or distances to, the stars associated with the cluster. The likelihoods of the observed parallaxes are far broader and the cluster is not obvious at all. In the posteriors there is a sharp peak in the probability at the previously determined distance to the cluster, the vertical lines, for far more stars than the likelihoods would imply. Our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy. Too strong?
